---
title: "TAIPAN REPORT"
author: "Aishwarya Anil Kumar, Arindom Baruah"
format: html
---


## Introduction

<!-- 
1. For your assigned Australian wild thing, read all you can about it, and write a paragraph describing it's key features, such as it's habitat, usual active time, whether numbers are increasing or declining, and even things like, if it is common for tourists to travel to see them, or there are special spotting promotions occurring as is happening [this month for the platypus](https://platypus.asn.au/report-a-sighting/). The write five unique expectations on what you might find when you explore the data, for example: 
<br><br>"We expect to see the occurrences only in the summer months."
<br>"We expect to find the most occurrences near the coast."
<br>"We expect to observe declining numbers since 1994."
<br><br> Be as specific as reasonably possible for your wild thing!
-->

## Data description

<!--
2. Download all the occurrence records from the Atlas of Living Australia, ideally scripted using the `atlas_occurrences()` function from the `galah` package. Make sure you include date, time, collector identification details. Save this into a `data-raw` folder in your repo.
-->
```{r}
#|warning: false
#|message: false
library(galah)
library(tidyverse)
```


```{r}

galah_config(email = "aani0005@student.monash.edu",
             download_reason_id = 10, 
             verbose = TRUE)
taipan <- atlas_occurrences(
  identify = galah_identify("Oxyuranus scutellatus scutellatus"))
save(taipan, file=here::here("data-raw/taipan_raw.rda"))
```

```{r}
glimpse(taipan)
```

### Description of the Variables in the Dataset

The dataset contains data on observations of the Coastal Taipan Snake (*Oxyuranus scutellatus scutellatus*). The dataset has been organized into a tidy format. Below is a detailed description of each variable:

| Variable Name    | Data Type | Description |
|------------------|-----------|-------------|
| Latitude         | dbl       | The latitude coordinate of the observation location. |
| Longitude        | dbl       | The longitude coordinate of the observation location. |
| eventDate        | date      | The date when the observation was made. |
| scientificName   | chr       | The scientific name of the observed species. In this dataset, it's *Oxyuranus scutellatus scutellatus*. |
| taxonConceptID   | chr       | A unique URL that provides information about the taxonomic concept for the species. |
| recordID         | chr       | A unique identifier for each observation record. |
| dataResourceName | chr       | The name of the data resource/provider. In this dataset, it is sourced from Queensland Museum for OZCAM. |
| occurrenceStatus | chr       | The status of the observation, indicating if the species is present or absent at the location. |


### Data Source and Cleaning Overview

**Data Source:** The data was sourced from the [Atlas of Living Australia (ALA)](https://www.ala.org.au/) using the `galah` R package.

**Data Cleaning Steps:**
1. Columns datatypes are changed for easement of further analysis such as the case with variable 'eventDate'.
2. Observations with missing event dates, longitude, or latitude were filtered out.


**Methods to Check and Clean Data:**

- **Validity Checks:** Ensure that the latitude and longitude values are within valid ranges. For latitude (-90 to 90) and for longitude (-180 to 180).
- **Consistency Checks:** Make sure there aren't any discrepancies in the way data is represented across rows.
- **Missing Value Analysis:** Any missing values should be analyzed to understand the reason behind them.
- **Outlier Analysis:** Especially for numeric fields like latitude and longitude, outlier detection can help spot any unusual or erroneous data points.
- **Date Consistency:** Ensure that the event dates fall within expected ranges and are formatted correctly.


### Citations:

1. **Atlas of Living Australia (ALA)** - https://www.ala.org.au/
2. **The `galah` package** - available on CRAN and also on GitHub. This package has been instrumental in fetching the data directly into the R environment. 

By exploring this data, researchers and enthusiasts can gain insights into the behavior, distribution, and occurrence patterns of the Coastal Taipan Snake, which can be useful for a range of ecological, environmental, and conservational studies.

## Initial data analysis

<!--
3. Perform a data quality check, and then go ahead and clean the data, saving only the wild sightings. Save the resulting data in a `data` folder as an `WILD_THING_COMMON_NAME.rda` file, so that any special variable types are clearly available. (Replace `WILD_THING_COMMON_NAME` with lower case common name of your assigned wild thing.) *Note:* Some are museum or historical records or live in an animal sanctuary or zoo, which do not truly capture their wild behaviour. Downloading the event information (`galah_select(group = "event")`) can help identify wild sightings.

Weather data pre-processing here too.
-->
```{r}
taipan <- taipan %>% 
  rename(Longitude = decimalLongitude,
         Latitude = decimalLatitude) %>%
  mutate(eventDate = as.Date(eventDate)) %>%
  filter(!is.na(eventDate)) %>%
  filter(!is.na(Longitude)) %>%
  filter(!is.na(Latitude))
save(taipan, file=here::here("data/taipan.rda"))

```


## Exploratory data analysis

<!--
4. Make plots of the data to check each of your expectations, and report what you learn. *Note:* This might also be a place where visual inference methods *may* be useful to determine if what is seen agrees or disputes what you expected. 

Weather data exploration here too.
-->

## Summary 

## References
